Elicitation Complexity via Convex Piecewise Linear Loss Functions

### 2019-01-13 notes-duality/bo-notes.tex
added accuracy condition to d-representation. bo@nyc

\begin{theorem}
  If an elicitable finite property is $d$-embeddable, then each of its level sets has a $d$-representation.
\end{theorem}
\begin{proof}
  Let $\ell_y: \reals^d \to \reals$ be the $y$th loss function.
  Write $\ell(r;p) = \sum_y p_y \ell_y(r)$ for the expected loss under report $r$.
  In particular, take $r$ to be the embedding point for a given level set.
  In this case, $r$ is optimal for belief $p$ if and only if $0 \in \partial \ell(r;p)$ where the subgradients are taken with respect to $r$ only.
  Therefore, the level set is exactly those $p$ such that $0 \in \partial \ell(r;p)$.
  We now construct these constraints from the loss functions.

@@

  Now, the subgradient set of a weighted sum is the weighted Minkowski sum of the subgradient sets:
    \[ \partial \ell(r;p) = \sum_y p_y \partial \ell_y(r) \]
  It follows that $\partial \ell(r;p)$ is the set of points $w$ satisfying that for all $v$, $\langle v,w \rangle \leq \sum_y p_y b_{vy}$.
  In particular therefore, $0 \in \partial \ell(r;p)$ if and only if $p$ is in the level set if and only if $0 \leq \sum_y p_y b_{vy}$ for all $v$.
  It is a polytope, so we can collect a finite set of $k$ vectors $v_1,\dots,v_k$ and form a matrix $B$ where $B_{iy} = b_{v_iy}$, and the level set is $\{p : B p \geq 0 \}$.
  Collect $v_1,\dots,v_k$ as the rows of a matrix $V$.
  For each $y$, the polytope $\partial \ell_y(r)$ satisfies the constraints $\langle v_i,w\rangle \leq B_{iw}$, i.e. is contained in $\{ w : Vw \leq B_y\}$.
  So that polytope is nonempty.
  So $V$ is a $d$-representation of the level set.
\end{proof}

We use this to prove lower bounds as follows:
\begin{enumerate}
  \item We're given a finite property.
  \item We pick any level set and write it as $\{ p : Bp \geq \vec{0}\}$ for a matrix $B \in \reals^{k \times n}$.
  \item We show there is no set of $k$ vectors $v_1,\dots,v_k$ that can form a $d$-representation.
  \item In particular, we show that however you pick $v_1,\dots,v_k$, for one of the outcomes $y$, the polytope $\{w : Vw \leq B_y \}$ is empty (infeasible).
\end{enumerate}

>>>>

\begin{definition}
  Say a $d$-representation $V$ of a polytope $\{p : Bp \geq \vec{0}\}$ is \emph{accurate} if for all $p$ in the polytope, $0$ is in the $p$-Minkowski combination:
  \[ 0 \in \sum_y p_y \{w : Vw \leq B_y \} . \]
\end{definition}

\begin{theorem}
  If an elicitable finite property is $d$-embeddable, then each of its level sets (1) has a $d$-representation that (2) is accurate.
\end{theorem}
\begin{proof}
  Let $\ell_y: \reals^d \to \reals$ be the $y$th loss function.
  Write $\ell(r;p) = \sum_y p_y \ell_y(r)$ for the expected loss under report $r$.
  In particular, take $r$ to be the embedding point for a given level set.
  In this case, $r$ is optimal for belief $p$ if and only if $0 \in \partial \ell(r;p)$ where the subgradients are taken with respect to $r$ only.
  Therefore, the level set is exactly those $p$ such that $0 \in \partial \ell(r;p)$.
  We now construct these constraints from the loss functions.

@@

  Now, the subgradient set of a weighted sum is the weighted Minkowski sum of the subgradient sets:
    \[ \partial \ell(r;p) = \sum_y p_y \partial \ell_y(r) \]
  It follows that $\partial \ell(r;p)$ is the set of points $w$ satisfying that for all $v$, $\langle v,w \rangle \leq \sum_y p_y b_{vy}$.
  In particular therefore, $0 \in \partial \ell(r;p)$ if and only if $p$ is in the level set if and only if $0 \leq \sum_y p_y b_{vy}$ for all $v$.
  It is a polytope, so we can collect a finite set of $k$ vectors $v_1,\dots,v_k$ and form a matrix $B$ where $B_{iy} = b_{v_iy}$, and the level set is $\{p : B p \geq 0 \}$.
  Collect $v_1,\dots,v_k$ as the rows of a matrix $V$.
  For each $y$, the polytope $\partial \ell_y(r)$ satisfies the constraints $\langle v_i,w\rangle \leq B_{iw}$, i.e. is contained in $P_y = \{ w : Vw \leq B_y\}$.
  So that polytope is nonempty.
  So $V$ is a $d$-representation of the level set.
  Furthermore, if $0 \in \partial \ell(r;p) = \sum_y p_y \partial \ell_y(r)$, then $0 \in \sum_y p_y P_y$, so $V$ is accurate.
\end{proof}

We use this to prove lower bounds as follows:
\begin{enumerate}
  \item We're given a finite property.
  \item We pick any level set and write it as $\{ p : Bp \geq \vec{0}\}$ for a matrix $B \in \reals^{k \times n}$.
  \item We show there is no set of $k$ vectors $v_1,\dots,v_k$ that can form a $d$-representation.
  \item In particular, we show that however you pick $v_1,\dots,v_k$, for one of the outcomes $y$, the polytope $\{w : Vw \leq B_y \}$ is empty (infeasible).
  \item Or, if it is possible that $d$-representations exist, we show that none of them can possibly be accuate.
\end{enumerate}
